---
title: "Functional Analysis"
output: html_notebook
---

```{r}
library(dplyr)
library(clusterProfiler)
library(ggplot2)
library(AnnotationForge)
library(DOSE)
library(enrichplot)
```

We are going to use AnnotationForge to create a database package that can be used with clusterProfiler. We will be using the makeOrgPackage() function. For this, we'll three dataframes:
gene_info - a df with gene ID (label as "GID") in the first column, "SYMBOL" in the second column, and "GENENAME" in the third.
chromosome - a df with "GID" in first column, "CHROMOSOME" (chromosome number) in second.
go - a df with "GID" in first column, "GO" (go term) in second column, and "EVIDENCE" in third column. Al GO term numbers will be unique (won't be repeated within column), but gene IDs can correspond to several different GO terms. 

Tutorial: https://www.bioconductor.org/packages/release/bioc/vignettes/AnnotationForge/inst/doc/MakingNewOrganismPackages.html


```{r}
# Load in GO annotated PA14 file, downloaded from Psuedomonas.com: https://www.pseudomonas.com/goterms/list?accession=&goterm=&goEvidenceCode=&ecoCode=&strain_id=109&term=Pseudomonas+aeruginosa+UCBPP-PA14&offset=0
PA14file <- read.csv("../data/GO_annotations/gene_ontology_csv.csv")
colnames(PA14file)

PA14file$chrom <- '1'

# Create gene_info df, remove duplicate rows
PA14Sym <- PA14file[,c("Locus.Tag", "Gene.Name", "Product.Description")]
PA14Sym <- PA14Sym[PA14Sym[,"Gene.Name"]!="-",]
PA14Sym <- PA14Sym[PA14Sym[,"Product.Description"]!="-",]
colnames(PA14Sym) <- c("GID","SYMBOL","GENENAME")
PA14Sym <- PA14Sym[!duplicated(PA14Sym), ]

# Create chromosome df, remove duplicate rows
PA14Chr <- PA14file[,c("Locus.Tag", "chrom")]
PA14Chr <- PA14Chr[PA14Chr[,"chrom"]!="-",]
colnames(PA14Chr) <- c("GID","CHROMOSOME")
PA14Chr %>% distinct(GID, .keep_all = TRUE)
PA14Chr <- PA14Chr[!duplicated(PA14Chr), ]

# Create go df
PA14Go <- PA14file[,c("Locus.Tag", "Accession", "GO.Evidence.Code")]
PA14Go <- PA14Go[PA14Go[,"GO.Evidence.Code"]!="",]
colnames(PA14Go) <- c("GID","GO","EVIDENCE")
PA14Go <- PA14Go[!duplicated(PA14Go), ]

# Call the function
makeOrgPackage(gene_info=PA14Sym, chromosome=PA14Chr, go=PA14Go,
               version="0.1",
               maintainer="Sabrina J Arif <sarif@umn.edu>",
               author="Sabrina J Arif <sarif@umn.edu>",
               outputDir = ".",
               tax_id="208963",
               genus="Pseudomonas",
               species=".aer.UCBPP.PA14",
               goTable="go")

# Install package
install.packages("./org.P.aer.UCBPP.PA14.eg.db", repos=NULL, type="source")
```

First let's prep some data for the enrichGO function. sig=genes of interest; uni=background genes.
```{r}
## Define genes of interest (sig)

# UPREGULATED ana4h vs unt4h
sig_upana4unt4 <- read.csv("../results/DESeq2/up_sigtab_ana4h_v_unt4h_alpha01beta1.csv")
sig_upana4unt4 <- as.character(sig_upana4unt4$GeneID)

# DOWNREGULATED ana4h vs unt4h
sig_downana4unt4 <- read.csv("../results/DESeq2/down_sigtab_ana4h_v_unt4h_alpha01beta1.csv")
sig_downana4unt4 <- as.character(sig_downana4unt4$GeneID)

# UPREGULATED hne4h vs unt4h
sig_uphne4unt4 <- read.csv("../results/DESeq2/up_sigtab_hne4h_v_unt4h_alpha01beta1.csv")
sig_uphne4unt4 <- as.character(sig_uphne4unt4$GeneID)

# DOWNREGULATED hne4h vs unt4h
sig_downhne4unt4 <- read.csv("../results/DESeq2/down_sigtab_hne4h_v_unt4h_alpha01beta1.csv")
sig_downhne4unt4 <- as.character(sig_downhne4unt4$GeneID)

# UPREGULATED ana16h vs unt16h
sig_upana16unt16 <- read.csv("../results/DESeq2/up_sigtab_ana16h_v_unt16h_alpha01beta1.csv")
sig_upana16unt16 <- as.character(sig_upana16unt16$GeneID)

# DOWNREGULATED ana16h vs unt16h
sig_downana16unt16 <- read.csv("../results/DESeq2/down_sigtab_ana16h_v_unt16h_alpha01beta1.csv")
sig_downana16unt16 <- as.character(sig_downana16unt16$GeneID)

# UPREGULATED ana16h vs ana4h
sig_upana16ana4 <- read.csv("../results/DESeq2/up_sigtab_ana16h_v_ana4h_alpha01beta1.csv")
sig_upana16ana4 <- as.character(sig_upana16ana4$GeneID)

# DOWNREGULATED ana16h vs unt16h
sig_downana16ana4 <- read.csv("../results/DESeq2/down_sigtab_ana16h_v_ana4h_alpha01beta1.csv")
sig_downana16ana4 <- as.character(sig_downana16ana4$GeneID)



### Define universes

## ana4h vs unt4h
# ALL
uni_ana4unt4 <- read.csv("../results/DESeq2/df_ana4h_v_unt4h.csv")
uni_ana4unt4 <- as.character(uni_ana4unt4$GeneID)
# SIG
uni_sigana4unt4 <- read.csv("../results/DESeq2/sigtab_ana4h_v_unt4h_alpha01beta1.csv")
uni_sigana4unt4 <- as.character(uni_sigana4unt4$GeneID)

## hne4h vs unt4h
# ALL
uni_hne4unt4 <- read.csv("../results/DESeq2/df_hne4h_v_unt4h.csv")
uni_hne4unt4 <- as.character(uni_hne4unt4$GeneID)
# SIG
uni_sighne4unt4 <- read.csv("../results/DESeq2/sigtab_hne4h_v_unt4h_alpha01beta1.csv")
uni_sighne4unt4 <- as.character(uni_sighne4unt4$GeneID)

## ana16h vs unt16h
# ALL
uni_ana16unt16 <- read.csv("../results/DESeq2/df_ana16h_v_unt16h.csv")
uni_ana16unt16 <- as.character(uni_ana16unt16$GeneID)

## ana16h vs ana4h
# ALL
uni_ana16ana4 <- read.csv("../results/DESeq2/df_ana16h_v_ana4h.csv")
uni_ana16ana4 <- as.character(uni_ana16ana4$GeneID)

# SIG
uni_sigana16unt16 <- read.csv("../results/DESeq2/sigtab_ana16h_v_unt16h_alpha01beta1.csv")
uni_sigana16unt16 <- as.character(uni_sigana16unt16$GeneID)

uni_sigana16ana4 <- read.csv("../results/DESeq2/sigtab_ana16h_v_ana4h_alpha01beta1.csv")
uni_sigana16ana4 <- as.character(uni_sigana16ana4$GeneID)


```

More prep for enrichment analysis
```{r}
# Load in the package you created
library(org.P.aer.UCBPP.PA14.eg.db)

# Create a function specified for the DESeq comparisons to reduce text
egoBP <- function(sig, uni){
  enrichGO(gene = sig, 
           universe = names(uni),
           keyType = "GID",
           OrgDb = org.P.aer.UCBPP.PA14.eg.db, 
           ont = "BP", 
           pAdjustMethod = "BH", 
           qvalueCutoff = 0.05, 
           readable = TRUE)}

# Function for converting GeneRatio to decimal and sorting in descending order
GO_tidy <- function(df){
  df$GeneRatioPerc <- sapply(df$GeneRatio, function(x) eval(parse(text=x))) %>%
  df[with(df, order(-GeneRatioPerc)), ]
}
```

Run enrichment analyses for the DESeq comparisons
```{r}
# UPREGULATED ana4h vs unt4h
egoBP_upana4unt4 <- egoBP(sig_upana4unt4, uni_ana4unt4)
resBP_upana4unt4 <- as.data.frame(egoBP_upana4unt4@result)
resBP_upana4unt4 <- resBP_upana4unt4[with(resBP_upana4unt4, order(-Count)), ]
#resBP_upana4unt4 <- GO_tidy(resBP_upana4unt4)
write.csv(resBP_upana4unt4, "../results/GO_enrichment/resBP_upana4unt4.csv", row.names=FALSE)

# DOWNREGULATED ana4h vs unt4h
egoBP_downana4unt4 <- egoBP(sig_downana4unt4, uni_ana4unt4)
resBP_downana4unt4 <- as.data.frame(egoBP_downana4unt4@result)
resBP_downana4unt4 <- resBP_downana4unt4[with(resBP_downana4unt4, order(-Count)), ]
write.csv(resBP_downana4unt4, "../results/GO_enrichment/resBP_downana4unt4.csv", row.names=FALSE)

# UPREGULATED hne4h vs unt4h
egoBP_uphne4unt4 <- egoBP(sig_uphne4unt4, uni_hne4unt4)
resBP_uphne4unt4 <- as.data.frame(egoBP_uphne4unt4@result)
write.csv(resBP_uphne4unt4, "../results/GO_enrichment/resBP_uphne4unt4.csv", row.names=FALSE)

# DOWNREGULATED hne4h vs unt4h
egoBP_downhne4unt4 <- egoBP(sig_downhne4unt4, uni_hne4unt4)
resBP_downhne4unt4 <- as.data.frame(egoBP_downhne4unt4@result)
write.csv(resBP_downhne4unt4, "../results/GO_enrichment/resBP_downhne4unt4.csv", row.names=FALSE)

# UPREGULATED ana16h vs unt16h
egoBP_upana16unt16 <- egoBP(sig_upana16unt16, uni_ana16unt16)
resBP_upana16unt16 <- as.data.frame(egoBP_upana16unt16@result)
resBP_upana16unt16 <- resBP_upana16unt16[with(resBP_upana16unt16, order(-Count)), ]
write.csv(resBP_upana16unt16, "../results/GO_enrichment/resBP_upana16unt16.csv", row.names=FALSE)

# DOWNREGULATED ana16h vs unt16h
egoBP_downana16unt16 <- egoBP(sig_downana16unt16, uni_ana16unt16)
resBP_downana16unt16 <- as.data.frame(egoBP_downana16unt16@result)
resBP_downana16unt16 <- resBP_downana16unt16[with(resBP_downana16unt16, order(-Count)), ]
write.csv(resBP_downana16unt16, "../results/GO_enrichment/resBP_downana16unt16.csv", row.names=FALSE)

# UPREGULATED ana16h vs unt16h
egoBP_upana16ana4 <- egoBP(sig_upana16ana4, uni_ana16ana4)
resBP_upana16ana4 <- as.data.frame(egoBP_upana16ana4@result)
resBP_upana16ana4 <- resBP_upana16ana4[with(resBP_upana16ana4, order(-Count)), ]
write.csv(resBP_upana16ana4, "../results/GO_enrichment/resBP_upana16ana4.csv", row.names=FALSE)

# DOWNREGULATED ana16h vs unt16h
egoBP_downana16ana4 <- egoBP(sig_downana16ana4, uni_ana16ana4)
resBP_downana16ana4 <- as.data.frame(egoBP_downana16ana4@result)
resBP_downana16ana4 <- resBP_downana16ana4[with(resBP_downana16ana4, order(-Count)), ]
write.csv(resBP_downana16ana4, "../results/GO_enrichment/resBP_downana16ana4.csv", row.names=FALSE)

# sig ana16h vs unt16h
egoBP_ana16ana4 <- egoBP(uni_sigana16ana4, uni_ana16ana4)
resBP_ana16ana4 <- as.data.frame(egoBP_ana16ana4@result)
resBP_ana16ana4 <- resBP_ana16ana4[with(resBP_ana16ana4, order(-Count)), ]
```

clusterprofiler dotplots
```{r}
dotplot(egoBP_upana4unt4, title="up_ana4h_v_unt4h / up_sig_n=130 / sig_n=275 / all_n=2905")
ggsave("../figures/Dotplots/dotplot_BP_upana4unt4.pdf", width=10, height=5)

dotplot(egoBP_downana4unt4, title="down_ana4h_v_unt4h / down_sig_n=145 / sig_n=275 / all_n=2905")
ggsave("../figures/Dotplots/dotplot_BP_downana4unt4.pdf", width=10, height=5)

dotplot(egoBP_uphne4unt4, title="up_hne4h_v_unt4h / up_sig_n=12 / sig_n=20 / all_n=2905")
ggsave("../figures/Dotplots/dotplot_BP_uphne4unt4.pdf", width=10, height=5)

dotplot(egoBP_downhne4unt4, title="down_hne4h_v_unt4h / down_sig_n=8 / sig_n=20 / all_n=2905")
ggsave("../figures/Dotplots/dotplot_BP_downhne4unt4.pdf", width=10, height=5)

dotplot(egoBP_upana16unt16, title="up_ana16h_v_unt16h / up_sig_n=32 / sig_n=122 / all_n=2905")
ggsave("../figures/Dotplots/dotplot_BP_upana16unt16.pdf", width=10, height=5)

dotplot(egoBP_downana16unt16, title="down_ana16h_v_unt16h / down_sig_n=90 / sig_n=122 / all_n=2905")
ggsave("../figures/Dotplots/dotplot_BP_downana16unt16.pdf", width=10, height=5)

dotplot(egoBP_upana16ana4, title="up_ana16h_v_ana4h")
ggsave("../figures/Dotplots/dotplot_BP_upana16ana4.pdf", width=10, height=5)

dotplot(egoBP_downana16ana4, title="down_ana16h_v_ana4h")
ggsave("../figures/Dotplots/dotplot_BP_downana16ana4.pdf", width=10, height=5)

```
```{r}
library(ggupset)
upsetplot(egoBP_upana16ana4, showCategory=20)
upsetplot(egoBP_downana16ana4, showCategory=20)

bp2 <- simplify(egoBP_upana16ana4, cutoff=0.9, by="p.adjust", select_fun=min)
bp3 <- simplify(egoBP_downana16ana4, cutoff=0.7, by="p.adjust", select_fun=min)

upsetplot(bp2, showCategory=20)
upsetplot(bp3, showCategory=20)
```
```{r}
## convert gene ID to Symbol
edox <- setReadable(egoBP_ana16ana4, 'org.P.aer.UCBPP.PA14.eg.db', 'GID')
p1 <- cnetplot(edox, foldChange=geneList)
## categorySize can be scaled by 'pvalue' or 'geneNum'
p2 <- cnetplot(edox, categorySize="pvalue", foldChange=geneList)
p3 <- cnetplot(edox, foldChange=geneList, circular = TRUE, colorEdge = TRUE)
cowplot::plot_grid(p1, p2, p3, ncol=3, labels=LETTERS[1:3], rel_widths=c(.8, .8, 1.2))
```
```{r}
p1 <- heatplot(edox)
p2 <- heatplot(edox, foldChange=geneList)
cowplot::plot_grid(p1, p2, ncol=1, labels=LETTERS[1:2])
```




